### 2019-07-08 methodology/novel-contributions.tex
Algorithmicx package

\begin{algorithm}[htp!]
	\caption{Exhaustive pseudo-MPE algorithm}
	\label{alg:pseudo-mpe-exhaustive}
	\begin{algorithmic}
		\STATE $evidence = $ user selected $(state,value)$ tuples
		\STATE $MPE\_tree = $ MPE Tree rooted in $evidence$
		\WHILE{True} 
			\STATE $mpe\_states$ = \texttt{next\_most\_probable\_states($evidence$)}
			\IF{$mpe\_states$ is not empty}
				\STATE $next\_state$ = head of $mpe\_states$ 
				\STATE propose $next\_state$ to user \COMMENT{the least entropic state}
				\IF{the user refuses $next\_state$}
					\FOR{$alternative\_state$ in $mpe\_states \smallsetminus next\_state$}
						\STATE propose $alternative\_state$ to user \COMMENT{the next least entropic states}
						\IF{the user accepts $alternative\_state$}
							\STATE call \texttt{generate\_alternative\_branch()} on $MPE\_tree$ 
							\STATE add $alternative\_state$ to $MPE\_tree$
							\STATE $evidence = evidence \cup alternative\_state$
						\ELSE
							\STATE continue
						\ENDIF
					\ENDFOR
				\ELSE
					\STATE add $next\_state$ to $MPE\_tree$
					\STATE $evidence = evidence \cup next\_state$
				\ENDIF
			\ELSE 
				\STATE return
			\ENDIF
		\ENDWHILE
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htp!]
	\caption{Independencies pseudo-MPE algorithm}
	\label{alg:pseudo-mpe-independencies}
	\begin{algorithmic}
		\STATE $evidence = $ user selected $(state,value)$ tuples
		\STATE $MPE\_tree = $ MPE Tree rooted in $evidence$
		\WHILE{True} 
			\STATE $separated = $ \texttt{evidence\_d\_separation($evidence$)} \COMMENT{based on evidence of previous step}
			\STATE $mpe\_states$ = \texttt{next\_most\_probable\_states($evidence$)}
			\STATE $mpe\_states = mpe\_states \smallsetminus separated$ 
			\IF{$mpe\_states$ is not empty}
				\STATE $next\_state$ = head of $mpe\_states$ 
				\STATE propose $next\_state$ to user \COMMENT{the least entropic state}
				\IF{the user refuses $next\_state$}
					\FOR{$alternative\_state$ in $mpe\_states \smallsetminus next\_state$}
						\STATE propose $alternative\_state$ to user \COMMENT{the next least entropic states}
						\IF{the user accepts $alternative\_state$}
							\STATE call \texttt{generate\_alternative\_branch()} on $MPE\_tree$ 
							\STATE add $alternative\_state$ to $MPE\_tree$
							\STATE $evidence = evidence \cup alternative\_state$
						\ELSE
							\STATE continue \COMMENT{go to next proposal}
						\ENDIF
					\ENDFOR
				\ELSE
					\STATE add $next\_state$ to $MPE\_tree$
					\STATE $evidence = evidence \cup next\_state$
				\ENDIF
			\ELSE 
				\STATE return 
			\ENDIF
		\ENDWHILE
	\end{algorithmic}
\end{algorithm}

@@

\todo{implementare e finire}
\begin{algorithm}[htp!]
	\caption{Thresholded pseudo-MPE algorithm}
	\label{alg:pseudo-mpe-thresholded}
	\begin{algorithmic}
		\STATE $separated\_list = \emptyset$
		\STATE {\bf input:} source $X$, evidence $E$, nodes $V$
		\FOR{target $Y \in V$} 
		\STATE append $d-separated(X, Y, E)$ to $separated\_list$ \COMMENT{will return true or false}
		\ENDFOR
		\STATE {\bf output:} $separated\_list$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[htp!]
	\caption{Exhaustive alternative explanation branch algorithm}
	\label{alg:alternative-branch-echaustive}
	\begin{algorithmic}
		\STATE $alternative\_evidence = evidence$ 
		\STATE $alt\_node = $ last node in the main MPE Tree
		\STATE $branch\_id = branch\_id + 1$
		\WHILE{True} 
			\STATE $mpe\_states$ = \texttt{next\_most\_probable\_states($alternative\_evidence$)}
			\IF{$mpe\_states$ is empty}
				\STATE return
			\ELSE
				\STATE $next\_state$ = head of $mpe\_states$ 
				\STATE create $next\_state$ node, tag it with $branch\_id$ and make it son of $alt\_node$
				\STATE update $alt\_node$ node to be $next\_state$ node
				\STATE $alternative\_evidence = alternative\_evidence \cup next\_state$
			\ENDIF
		\ENDWHILE
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htp!]
	\caption{Independencies alternative explanation branch algorithm}
	\label{alg:alternative-branch-independencies}
	\begin{algorithmic}
		\STATE $alternative\_evidence = evidence$ 
		\STATE $alt\_node = $ last node in the main MPE Tree
		\STATE $branch\_id = branch\_id + 1$
		\WHILE{True} 
			\STATE $separated = $ \texttt{evidence\_d\_separation($alternative\_evidence$)}
			\STATE $mpe\_states$ = \texttt{next\_most\_probable\_states($alternative\_evidence$)}
			\STATE $mpe\_states = mpe\_states \smallsetminus separated$ 
			\IF{$mpe\_states$ is empty}
				\STATE return
			\ELSE
				\STATE $next\_state$ = head of $mpe\_states$ 
				\STATE create $next\_state$ node, tag it with $branch\_id$ and make it son of $alt\_node$
				\STATE update $alt\_node$ node to be $next\_state$ node
				\STATE $alternative\_evidence = alternative\_evidence \cup next\_state$
			\ENDIF
		\ENDWHILE
	\end{algorithmic}
\end{algorithm}

\todo{implementare e fare}
\begin{algorithm}[htp!]
	\caption{Independencies alternative explanation branch algorithm}
	\label{alg:alternative-branch-thresholded}
	\begin{algorithmic}
		\STATE $evidence = $ user selected $(state,value)$ tuples
		\STATE $MPE\_tree = $ root MPE Tree in $evidence$
		\WHILE{True} 
			\STATE $separated = $ \texttt{evidence\_d\_separation()} \COMMENT{based on evidence of previous step}
			\STATE $mpe\_states$ = \texttt{next\_most\_probable\_states()}
			\STATE $mpe\_states = mpe\_states \smallsetminus separated$ 
			\IF{$mpe\_states$ is not empty}
				\STATE $next\_state$ = head of $mpe\_states$ 
				\STATE propose $next\_state$ to user \COMMENT{the least entropic state}
				\IF{the user refuses $next\_state$}
					\FOR{$alternative\_state$ in $mpe\_states \smallsetminus next\_state$}
						\STATE propose $alternative\_state$ to user \COMMENT{the next least entropic states}
						\IF{the user accepts $alternative\_state$}
							\STATE call \texttt{generate\_alternative\_branch()} on $MPE\_tree$ 
							\STATE add $alternative\_state$ to $MPE\_tree$
							\STATE $evidence = evidence \cup alternative\_state$
						\ELSE
							\STATE continue \COMMENT{go to next proposal}
						\ENDIF
					\ENDFOR
				\ELSE
					\STATE add $next\_state$ to $MPE\_tree$
					\STATE $evidence = evidence \cup next\_state$
				\ENDIF
			\ELSE 
				\STATE return 
			\ENDIF
		\ENDWHILE
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[htp!]
	\caption{Pseudo-MPE from random evidence algorithm}
	\label{alg:pseudo-mpe-random-evidence}
	\begin{algorithmic}
		\STATE $evidence = \emptyset$
		\STATE generate random number $k \in [ 1, |V| ]$
		\STATE $S = $ choose $k$ variables from $V$
		\FOR{$s \in S$}
			\STATE choose random $v$ in the possible values of $e$ 
			\STATE $evidence = evidence \cup (s,v)$
		\ENDFOR
		\STATE $MPE\_tree = $ MPE Tree rooted in $evidence$
		\STATE $last\_node = evidence$
		\STATE $alternative\_evidence = evidence$ 
		\WHILE{True} 
			\STATE $mpe\_states$ = \texttt{next\_most\_probable\_states($alternative\_evidence$)}
			\IF{$mpe\_states$ is empty}
				\STATE return
			\ELSE
				\STATE $next\_state$ = head of $mpe\_states$ 
				\STATE create $next\_state$ node and make it son of $last\_node$
				\STATE update $alt\_node$ node to be $next\_state$ node
				\STATE $alternative\_evidence = alternative\_evidence \cup next\_state$
			\ENDIF
		\ENDWHILE
	\end{algorithmic}
\end{algorithm}

>>>>

\begin{algorithm}[htp!]
	\caption{Exhaustive pseudo-MPE algorithm}
	\label{alg:pseudo-mpe-exhaustive}
	\begin{algorithmic}[1]
		\State $evidence = $ user selected $(state,value)$ tuples
		\State $MPE\_tree = $ MPE Tree rooted in $evidence$
		\While{True} 
			\State $mpe\_states$ = \texttt{next\_most\_probable\_states($evidence$)}
			\If{$mpe\_states$ is not empty}
				\State $next\_state$ = head of $mpe\_states$ 
				\State propose $next\_state$ to user \Comment{the least entropic state}
				\If{the user refuses $next\_state$}
					\For{$alternative\_state$ in $mpe\_states \smallsetminus next\_state$}
						\State propose $alternative\_state$ to user \Comment{the next least entropic states}
						\If{the user accepts $alternative\_state$}
							\State call \texttt{generate\_alternative\_branch()} on $MPE\_tree$ 
							\State add $alternative\_state$ to $MPE\_tree$
							\State $evidence = evidence \cup alternative\_state$
						\Else
							\State continue
						\EndIf
					\EndFor
				\Else
					\State add $next\_state$ to $MPE\_tree$
					\State $evidence = evidence \cup next\_state$
				\EndIf
			\Else 
				\State return
			\EndIf
		\EndWhile
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htp!]
	\caption{Independencies pseudo-MPE algorithm}
	\label{alg:pseudo-mpe-independencies}
	\begin{algorithmic}[1]
		\State $evidence = $ user selected $(state,value)$ tuples
		\State $MPE\_tree = $ MPE Tree rooted in $evidence$
		\While{True} 
			\State $separated = $ \texttt{evidence\_d\_separation($evidence$)} \Comment{based on evidence of previous step}
			\State $mpe\_states$ = \texttt{next\_most\_probable\_states($evidence$)}
			\State $mpe\_states = mpe\_states \smallsetminus separated$ 
			\If{$mpe\_states$ is not empty}
				\State $next\_state$ = head of $mpe\_states$ 
				\State propose $next\_state$ to user \Comment{the least entropic state}
				\If{the user refuses $next\_state$}
					\For{$alternative\_state$ in $mpe\_states \smallsetminus next\_state$}
						\State propose $alternative\_state$ to user \Comment{the next least entropic states}
						\If{the user accepts $alternative\_state$}
							\State call \texttt{generate\_alternative\_branch()} on $MPE\_tree$ 
							\State add $alternative\_state$ to $MPE\_tree$
							\State $evidence = evidence \cup alternative\_state$
						\Else
							\State continue \Comment{go to next proposal}
						\EndIf
					\EndFor
				\Else
					\State add $next\_state$ to $MPE\_tree$
					\State $evidence = evidence \cup next\_state$
				\EndIf
			\Else 
				\State return 
			\EndIf
		\EndWhile
	\end{algorithmic}
\end{algorithm}

@@

\todo{implementare e finire}
\begin{algorithm}[htp!]
	\caption{Thresholded pseudo-MPE algorithm}
	\label{alg:pseudo-mpe-thresholded}
	\begin{algorithmic}[1]
		\State $separated\_list = \emptyset$
		\State {\bf input:} source $X$, evidence $E$, nodes $V$
		\For{target $Y \in V$} 
		\State append $d-separated(X, Y, E)$ to $separated\_list$ \Comment{will return true or false}
		\EndFor
		\State {\bf output:} $separated\_list$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[htp!]
	\caption{Exhaustive alternative explanation branch algorithm}
	\label{alg:alternative-branch-echaustive}
	\begin{algorithmic}[1]
		\State $alternative\_evidence = evidence$ 
		\State $alt\_node = $ last node in the main MPE Tree
		\State $branch\_id = branch\_id + 1$
		\While{True} 
			\State $mpe\_states$ = \texttt{next\_most\_probable\_states($alternative\_evidence$)}
			\If{$mpe\_states$ is empty}
				\State return
			\Else
				\State $next\_state$ = head of $mpe\_states$ 
				\State create $next\_state$ node, tag it with $branch\_id$ and make it son of $alt\_node$
				\State update $alt\_node$ node to be $next\_state$ node
				\State $alternative\_evidence = alternative\_evidence \cup next\_state$
			\EndIf
		\EndWhile
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htp!]
	\caption{Independencies alternative explanation branch algorithm}
	\label{alg:alternative-branch-independencies}
	\begin{algorithmic}[1]
		\State $alternative\_evidence = evidence$ 
		\State $alt\_node = $ last node in the main MPE Tree
		\State $branch\_id = branch\_id + 1$
		\While{True} 
			\State $separated = $ \texttt{evidence\_d\_separation($alternative\_evidence$)}
			\State $mpe\_states$ = \texttt{next\_most\_probable\_states($alternative\_evidence$)}
			\State $mpe\_states = mpe\_states \smallsetminus separated$ 
			\If{$mpe\_states$ is empty}
				\State return
			\Else
				\State $next\_state$ = head of $mpe\_states$ 
				\State create $next\_state$ node, tag it with $branch\_id$ and make it son of $alt\_node$
				\State update $alt\_node$ node to be $next\_state$ node
				\State $alternative\_evidence = alternative\_evidence \cup next\_state$
			\EndIf
		\EndWhile
	\end{algorithmic}
\end{algorithm}

\todo{implementare e fare}
\begin{algorithm}[htp!]
	\caption{Independencies alternative explanation branch algorithm}
	\label{alg:alternative-branch-thresholded}
	\begin{algorithmic}[1]
		\State $evidence = $ user selected $(state,value)$ tuples
		\State $MPE\_tree = $ root MPE Tree in $evidence$
		\While{True} 
			\State $separated = $ \texttt{evidence\_d\_separation()} \Comment{based on evidence of previous step}
			\State $mpe\_states$ = \texttt{next\_most\_probable\_states()}
			\State $mpe\_states = mpe\_states \smallsetminus separated$ 
			\If{$mpe\_states$ is not empty}
				\State $next\_state$ = head of $mpe\_states$ 
				\State propose $next\_state$ to user \Comment{the least entropic state}
				\If{the user refuses $next\_state$}
					\For{$alternative\_state$ in $mpe\_states \smallsetminus next\_state$}
						\State propose $alternative\_state$ to user \Comment{the next least entropic states}
						\If{the user accepts $alternative\_state$}
							\State call \texttt{generate\_alternative\_branch()} on $MPE\_tree$ 
							\State add $alternative\_state$ to $MPE\_tree$
							\State $evidence = evidence \cup alternative\_state$
						\Else
							\State continue \Comment{go to next proposal}
						\EndIf
					\EndFor
				\Else
					\State add $next\_state$ to $MPE\_tree$
					\State $evidence = evidence \cup next\_state$
				\EndIf
			\Else 
				\State return 
			\EndIf
		\EndWhile
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[htp!]
	\caption{Pseudo-MPE from random evidence algorithm}
	\label{alg:pseudo-mpe-random-evidence}
	\begin{algorithmic}[1]
		\State $evidence = \emptyset$
		\State generate random number $k \in [ 1, |V| ]$
		\State $S = $ choose $k$ variables from $V$
		\For{$s \in S$}
			\State choose random $v$ in the possible values of $e$ 
			\State $evidence = evidence \cup (s,v)$
		\EndFor
		\State $MPE\_tree = $ MPE Tree rooted in $evidence$
		\State $last\_node = evidence$
		\State $alternative\_evidence = evidence$ 
		\While{True} 
			\State $mpe\_states$ = \texttt{next\_most\_probable\_states($alternative\_evidence$)}
			\If{$mpe\_states$ is empty}
				\State return
			\Else
				\State $next\_state$ = head of $mpe\_states$ 
				\State create $next\_state$ node and make it son of $last\_node$
				\State update $alt\_node$ node to be $next\_state$ node
				\State $alternative\_evidence = alternative\_evidence \cup next\_state$
			\EndIf
		\EndWhile
	\end{algorithmic}
\end{algorithm}
